Explore Vegetation Data

### 2016-01-07 explore_veg_data.Rmd
Recalculate/correct alt species abundances.

alt_sp_live <- filter(alt_sp_dbh, Dead == 0)
alt_sp_dead <- filter(alt_sp_dbh, Dead == 1)

alt_sp_summ <- left_join(
      # First data frame
      alt_sp_live %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
      summarise(live_count = length(Species), live_avg_dbh = mean(dbh), live_tot_dbh = sum(dbh)),
      # Second data frame
      alt_sp_dead %>% filter(year != 2012) %>% group_by(PlotID, Species, year) %>% 
      summarise(dead_count = length(Species), dead_avg_dbh = mean(dbh), dead_tot_dbh = sum(dbh))
)
summary(alt_sp_summ)

write.csv(alt_sp_summ, "analysis/data/untagged_dbh_summary.csv")
```

>>>>

summary(filter(alt_sp_dbh, year == 2003))
unique(droplevels(filter(alt_sp_dbh, year == 2003)$PlotID))
pid2003 <- as.character(unique(filter(alt_sp_dbh, year == 2003)$PlotID))
alt_sp_2003 <- filter(alt_sp_dbh, year == 2003)

summary(filter(alt_sp_dbh, year == 2004))
unique(droplevels(filter(alt_sp_dbh, year == 2004)$PlotID))
pid2004 <- as.character(unique(filter(alt_sp_dbh, year == 2004)$PlotID))
alt_sp_2004 <- filter(alt_sp_dbh, year == 2004)

pid2003 %in% pid2004
pid2004[47]; pid2004[69]
filter(alt_sp_dbh, year == 2003, PlotID == "MITSU02")
filter(alt_sp_dbh, year == 2004, PlotID == "MITSU02")

summary(filter(alt_sp_dbh, year == 2005))
unique(droplevels(filter(alt_sp_dbh, year == 2005)$PlotID))
pid2005 <- as.character(unique(filter(alt_sp_dbh, year == 2005)$PlotID))
alt_sp_2005 <- filter(alt_sp_dbh, year == 2005)

# I think I should work this backwards so that the most recent sampling is added to
pid2005 %in% pid2004# COOK02 is the only duplicate in the 2004 sampling
filter(alt_sp_2004, PlotID=="COOK02")
filter(alt_sp_2005, PlotID=="COOK02")
# But there are a lot more records in the 2004 data, which is closer to the 2012 & 2014 sampling. I could combine the 2004 & 2005 data, including the potentially duplicated COOK02 data.
summary(filter(alt_sp_2012, PlotID=="COOK02"))
filter(alt_sp_2014, PlotID=="COOK02")

alt_sp_45 <- rbind(alt_sp_2005, alt_sp_2004)
summary(alt_sp_45)
pid45 <- as.character(unique(alt_sp_45$PlotID))# 144 unique plots
pid45 %in% pid2003
pid45[71]; pid45[93]; pid45[133]
filter(alt_sp_45, PlotID == "MITSU02")
filter(alt_sp_2003, PlotID == "MITSU02")
# Exclude MITSU02 from the 2003 data
alt_sp_2003 <- filter(alt_sp_2003, PlotID != "MITSU02")

filter(alt_sp_45, PlotID == "LINS01")
filter(alt_sp_2003, PlotID == "LINS01")
# Keep LINS01 from 2003 data, add to other data
filter(alt_sp_45, PlotID == "GANAY02")
filter(alt_sp_2003, PlotID == "GANAY02")
# Exclude GANAY02 from the 2004, keep the 2003 data
alt_sp_45 <- filter(alt_sp_45, PlotID != "GANAY02")
unique(alt_sp_2003$PlotID) %in% unique(alt_sp_45$PlotID)
alt_sp_45 <- rbind(alt_sp_45, alt_sp_2003)
alt_sp_2005 <- alt_sp_45
rm(alt_sp_45)

summary(filter(alt_sp_dbh, year == 2012))
pid2012 <- unique(droplevels(filter(alt_sp_dbh, year == 2012)$PlotID))# 171 plots
alt_sp_2012 <- filter(alt_sp_dbh, year == 2012)

summary(filter(alt_sp_dbh, year == 2014))
pid2014 <- unique(droplevels(filter(alt_sp_dbh, year == 2014)$PlotID))# 164 plots
alt_sp_2014 <- filter(alt_sp_dbh, year == 2014)

# Determine which plots from 2012 (the one with more records) aren't in the 2014 data.
pid2012 %in% pid2014
pid2012sub <- droplevels(pid2012[c(4,12:13,20:21,52:53,111,114)])
alt_sp_2012 <- filter(alt_sp_2012, PlotID == "ARBIT04" | PlotID == "LINS01" | PlotID == "LUPIN01" | PlotID == "MITSU02" | PlotID == "MITSU04" | PlotID == "SKILE01" | PlotID == "SPAUL01" | PlotID == "GANAY02" | PlotID == "HALOW02")

alt_sp_2014 <- rbind(alt_sp_2014, alt_sp_2012)
summary(alt_sp_2014)
summary(alt_sp_2005)
alt_sp_2005$year <- 2005

alt_sp_2014 <- select(alt_sp_2014, -date)
alt_sp_2005 <- select(alt_sp_2005, -date)

# Create data frames of live and dead stems for '2005' and '2014' samplings. In reality the 2005 grouping has data from visits during 2003 and 2004, and the 2014 grouping includes data from some visits in 2012.
alt_sp_2014_live <- filter(alt_sp_2014, Dead == 0)
alt_sp_2014_live <- select(alt_sp_2014_live, -Dead)
alt_sp_2005_live <- filter(alt_sp_2005, Dead == 0)
alt_sp_2005_live <- select(alt_sp_2005_live, -Dead)
alt_sp_2014_dead <- filter(alt_sp_2014, Dead == 1)
alt_sp_2014_dead <- select(alt_sp_2014_dead, -Dead)
alt_sp_2005_dead <- filter(alt_sp_2005, Dead == 1)
alt_sp_2005_dead <- select(alt_sp_2005_dead, -Dead)

# Confirm year assignments are correct/consistent.
unique(alt_sp_2014_live$year)
unique(alt_sp_2005_live$year)
unique(alt_sp_2005_dead$year)
unique(alt_sp_2014_dead$year)

alt_sp_summ_2005 <- left_join(
      # First data frame
      alt_sp_2005_live %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  live_count = length(Species),
                  live_avg_dbh = mean(dbh),
                  live_tot_dbh = sum(dbh)),
      # Second data frame
      alt_sp_2005_dead %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  dead_count = length(Species), 
                  dead_avg_dbh = mean(dbh), 
                  dead_tot_dbh = sum(dbh))
)
summary(alt_sp_summ_2005)

alt_sp_summ_2014 <- left_join(
      # First data frame
      alt_sp_2014_live %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  live_count = length(Species),
                  live_avg_dbh = mean(dbh),
                  live_tot_dbh = sum(dbh)),
      # Second data frame
      alt_sp_2014_dead %>%
            group_by(PlotID, Species, year) %>% 
            summarise(
                  dead_count = length(Species), 
                  dead_avg_dbh = mean(dbh), 
                  dead_tot_dbh = sum(dbh))
)
summary(alt_sp_summ_2014)

# Write out summary data frames of the untagged DBH species
write.csv(alt_sp_summ_2005, "analysis/data/untagged_dbh2005_summary.csv", row.names = F)
write.csv(alt_sp_summ_2014, "analysis/data/untagged_dbh2014_summary.csv", row.names = F)
```

```{r make wide format species abundance}
unique(alt_sp_2014_live$Species)
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="arcto sp."] <- "arcto"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="prunus sp."] <- "prunus"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="unknown sp. 1"] <- "unk1"
alt_sp_2014_live$Species[alt_sp_2014_live$Species=="quercus sp."] <- "qusp"

unique(alt_sp_2005_live$Species)
alt_sp_2005_live$Species[alt_sp_2005_live$Species=="quercus sp."] <- "qusp"
alt_sp_2005_live$Species[alt_sp_2005_live$Species=="arcto sp."] <- "arcto"
alt_sp_2005_live$Species[alt_sp_2005_live$Species=="prunus sp."] <- "prunus"

# Aggregate to species counts by plot
alt_sp_05_live_abund <- alt_sp_2005_live %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
# Spread from long to wide format
alt_sp_05_live_abund <- alt_sp_05_live_abund %>% spread(Species, count)

alt_sp_14_live_abund <- alt_sp_2014_live %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
alt_sp_14_live_abund <- alt_sp_14_live_abund %>% spread(Species, count)

# Create species abundance data frames of dead species
alt_sp_05_dead_abund <- alt_sp_2005_dead %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
alt_sp_05_dead_abund <- alt_sp_05_dead_abund %>% spread(Species, count)

alt_sp_14_dead_abund <- alt_sp_2014_dead %>% select(-dbh, -year) %>% 
      group_by(PlotID, Species) %>% summarise(count = length(Species))
alt_sp_14_dead_abund <- alt_sp_14_dead_abund %>% spread(Species, count)

# Write out wide format abundance data frames
write.csv(alt_sp_05_live_abund, "analysis/data/altsp05abund_live.csv")
write.csv(alt_sp_14_live_abund, "analysis/data/altsp14abund_live.csv")
write.csv(alt_sp_05_dead_abund, "analysis/data/altsp05abund_dead.csv")
write.csv(alt_sp_14_dead_abund, "analysis/data/altsp14abund_dead.csv")
```
